---
Title: "Homework 2"
Subtitle: "Submission 2"
Author: "Alekhya Pidugu"
format:
  pdf:
    output-file: "pidugu-a-hwk1-2"
    output-ext:  "pdf"
---
# Git Repository: https://github.com/AlekhyaPidugu/Homework2

```{r results="hide", echo=FALSE, message=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes, fixest,
               scales, gganimate, gapminder, gifski, png, tufte, plotly, OECD,
               ggrepel, survey, foreign, devtools, pdftools, kableExtra, modelsummary,
               kableExtra)

```

\newpage
# Question 1
```{r results="hide", echo=FALSE, message=FALSE}
final.hcris.data <- read_rds("/Users/alekhyapidugu/Desktop/Homework2/data/output/HCRIS_Data.rds")
final.hcris.v1996=read_rds('/Users/alekhyapidugu/Desktop/Homework2/data/output/HCRIS_Data_v1996.rds')
final.hcris.v2010=read_rds('/Users/alekhyapidugu/Desktop/Homework2/data/output/HCRIS_Data_v2010.rds')

## create missing variables for columns introduced in v2010 of hcris forms
final.hcris.v1996 = final.hcris.v1996 %>%
  mutate(hvbp_payment=NA, hrrp_payment=NA)

## combine v1996 and v2010 hcris forms, and sort by provider_number/year
final.hcris=rbind(final.hcris.v1996,final.hcris.v2010) %>%
  mutate(fy_end=mdy(fy_end),fy_start=mdy(fy_start),
         date_processed=mdy(date_processed),date_created=mdy(date_created),
         tot_discounts=abs(tot_discounts), hrrp_payment=abs(hrrp_payment)) %>%
  mutate(fyear=year(fy_end)) %>%
  arrange(provider_number,fyear) %>%
  select(-year)

## count of hospitals/provider_number by year
final.hcris %>% group_by(fyear) %>% count()

# Clean data --------------------------------------------------------------

## create count of reports by hospital fiscal year
final.hcris =
  final.hcris %>% 
  add_count(provider_number, fyear, name="total_reports")

## create running total of reports
final.hcris =
  final.hcris %>% 
  group_by(provider_number, fyear) %>%
  mutate(report_number=row_number())

## identify hospitals with only one report per fiscal year 
## this will be the first set of hospitals in the final dataset
unique.hcris1 =
  final.hcris %>%
  filter(total_reports==1) %>%
  select(-report, -total_reports, -report_number, -npi, -status) %>%
  mutate(source='unique reports')


## identify hospitals with multiple reports per fiscal year
duplicate.hcris = 
  final.hcris %>%
  filter(total_reports>1) %>%
  mutate(time_diff=fy_end-fy_start)

library(ggplot2)

# Plot the number of hospitals with multiple reports over time
hospitals_over_time <- duplicate.hcris %>%
  group_by(fyear) %>%
  summarise(num_hospitals = n())

ggplot(hospitals_over_time, aes(x = fyear, y = num_hospitals)) +
  geom_line() +
  labs(x = "Fiscal Year", y = "Number of Hospitals with Multiple Reports") +
  ggtitle("Hospitals with Multiple Reports in the Same Year")
```

\newpage
# Question 2
<br>
There are 6486 unique hospital IDs.

\newpage
# Question 3
```{r results="hide", echo=FALSE, message=FALSE}
HCRIS_Data <- read_rds("/Users/alekhyapidugu/Desktop/Homework2/data/output/HCRIS_Data.rds")
HCRIS_Data_v1996=read_rds('/Users/alekhyapidugu/Desktop/Homework2/data/output/HCRIS_Data_v1996.rds')
HCRIS_Data_v2010=read_rds('/Users/alekhyapidugu/Desktop/Homework2/data/output/HCRIS_Data_v2010.rds')
# Select relevant columns for analysis
selected_columns <- c("year", "tot_charges")
charges_data_v1996 <- HCRIS_Data_v1996[, selected_columns]
charges_data_v2010 <- HCRIS_Data_v2010[, selected_columns]

# Add a dataset identifier for each year
charges_data_v1996$dataset <- "1996"
charges_data_v2010$dataset <- "2010"

# Combine the data frames
charges_data <- rbind(charges_data_v1996, charges_data_v2010)

# Filter out rows with missing or zero charges
charges_data <- charges_data %>%
  filter(!is.na(tot_charges) & tot_charges > 0)

# Create a violin plot
ggplot(charges_data, aes(x = factor(year), y = tot_charges, fill = factor(year))) +
  geom_violin() +
  labs(title = "Distribution of Total Charges by Year",
       x = "Year",
       y = "Total Charges",
       fill = "Year") +
  theme_minimal()
  ```

\newpage
# Question 4
```{r results="hide", echo=FALSE, message=FALSE}
HCRIS_Data <- read_rds("/Users/alekhyapidugu/Desktop/Homework2/data/output/HCRIS_Data.rds")
HCRIS_Data_v1996=read_rds('/Users/alekhyapidugu/Desktop/Homework2/data/output/HCRIS_Data_v1996.rds')
HCRIS_Data_v2010=read_rds('/Users/alekhyapidugu/Desktop/Homework2/data/output/HCRIS_Data_v2010.rds')
# Select relevant columns for analysis
selected_columns <- c("year", "tot_discounts", "tot_charges", "ip_charges", "icu_charges", "ancillary_charges",
                       "tot_mcare_payment", "tot_discharges", "mcare_discharges")

# Extract necessary columns from both datasets
price_data_v1996 <- HCRIS_Data_v1996[, selected_columns]
price_data_v2010 <- HCRIS_Data_v2010[, selected_columns]

# Combine the datasets
price_data <- rbind(
  mutate(price_data_v1996, dataset = "1996"),
  mutate(price_data_v2010, dataset = "2010")
)

# Calculate discount factor and estimated prices
price_data <- price_data %>%
  mutate(discount_factor = 1 - tot_discounts / tot_charges,
         price_num = (ip_charges + icu_charges + ancillary_charges) * discount_factor - tot_mcare_payment,
         price_denom = tot_discharges - mcare_discharges,
         estimated_price = ifelse(price_denom > 0, price_num / price_denom, NA_real_))

# Filter out outliers and negative prices
price_data <- price_data %>%
  filter(!is.na(estimated_price) & estimated_price > 0)

#Create Violin plot
ggplot(price_data, aes(x = factor(year), y = estimated_price, fill = factor(year))) +
  geom_violin() +
  labs(title = "Distribution of Estimated Prices by Year",
       x = "Year",
       y = "Estimated Prices",
       fill = "Year") +
  theme_minimal()
  ```

\newpage
# Question 5
<br> Not Penalized    9551    
<br> Penalized        9883 
<br>

\newpage
# Question 6
<br> 0      7612       7465      8706       8287  
 <br>1      6981       8529      8584       8566  
 <br>

\newpage
# Question 7
```{r echo=FALSE, message=FALSE}
ate_values <- c(199.5281, 199.5281, 199.5281, 199.5281)
estimators <- c("Nearest Neighbor (Inverse Variance Distance)",
                "Nearest Neighbor (Mahalanobis Distance)",
                "Inverse Propensity Weighting",
                "Simple Linear Regression")

# Create a data frame
ate_table <- data.frame(Estimator = estimators, ATE = ate_values)

# Print the table using kable
knitr::kable(ate_table, caption = "ATE values for different estimators", format = "markdown")

```

\newpage
# Question 8 
<br>
The results from the various treatment effect estimators are the same. 
<br>

\newpage
# Question 9 
<br>
I think it is hard to claim that there is a causal effect of the penalty with a controlled design to test this. But using just observation with this data there may be a causal effect but you just have to make sure you account for confounding factors when making this interpretation of the data.
<br>

\newpage
# Question 10 
<br>
I take back what I said about this data being easier to work with. Quarto is still very challenging but I got the PDF to work. I could not load my work space becuase it kept saying the directory did not exist. But I copied the file path directly so I do not know. Finally figured out question 7. 