---
Title: "Homework 2"
Subtitle: "Submission 2"
Author: "Alekhya Pidugu"
format:
  pdf:
    output-file: "pidugu-a-hwk1-2"
    output-ext:  "pdf"
---
# Git Repository: 

```{r results="hide", echo=FALSE, message=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes, fixest,
               scales, gganimate, gapminder, gifski, png, tufte, plotly, OECD,
               ggrepel, survey, foreign, devtools, pdftools, kableExtra, modelsummary,
               kableExtra)

```

\newpage
# Question 1
```{r results="hide", echo=FALSE, message=FALSE}
final.hcris.data <- read_rds("/Users/alekhyapidugu/Desktop/Homework2/data/output/HCRIS_Data.rds")
final.hcris.v1996=read_rds('/Users/alekhyapidugu/Desktop/Homework2/data/output/HCRIS_Data_v1996.rds')
final.hcris.v2010=read_rds('/Users/alekhyapidugu/Desktop/Homework2/data/output/HCRIS_Data_v2010.rds')

## create missing variables for columns introduced in v2010 of hcris forms
final.hcris.v1996 = final.hcris.v1996 %>%
  mutate(hvbp_payment=NA, hrrp_payment=NA)

## combine v1996 and v2010 hcris forms, and sort by provider_number/year
final.hcris=rbind(final.hcris.v1996,final.hcris.v2010) %>%
  mutate(fy_end=mdy(fy_end),fy_start=mdy(fy_start),
         date_processed=mdy(date_processed),date_created=mdy(date_created),
         tot_discounts=abs(tot_discounts), hrrp_payment=abs(hrrp_payment)) %>%
  mutate(fyear=year(fy_end)) %>%
  arrange(provider_number,fyear) %>%
  select(-year)

## count of hospitals/provider_number by year
final.hcris %>% group_by(fyear) %>% count()

# Clean data --------------------------------------------------------------

## create count of reports by hospital fiscal year
final.hcris =
  final.hcris %>% 
  add_count(provider_number, fyear, name="total_reports")

## create running total of reports
final.hcris =
  final.hcris %>% 
  group_by(provider_number, fyear) %>%
  mutate(report_number=row_number())

## identify hospitals with only one report per fiscal year 
## this will be the first set of hospitals in the final dataset
unique.hcris1 =
  final.hcris %>%
  filter(total_reports==1) %>%
  select(-report, -total_reports, -report_number, -npi, -status) %>%
  mutate(source='unique reports')


## identify hospitals with multiple reports per fiscal year
duplicate.hcris = 
  final.hcris %>%
  filter(total_reports>1) %>%
  mutate(time_diff=fy_end-fy_start)

library(ggplot2)

# Plot the number of hospitals with multiple reports over time
hospitals_over_time <- duplicate.hcris %>%
  group_by(fyear) %>%
  summarise(num_hospitals = n())

ggplot(hospitals_over_time, aes(x = fyear, y = num_hospitals)) +
  geom_line() +
  labs(x = "Fiscal Year", y = "Number of Hospitals with Multiple Reports") +
  ggtitle("Hospitals with Multiple Reports in the Same Year")
```

\newpage
# Question 2
<br>
There are 6486 unique hospital IDs.

\newpage
# Question 3
```{r results="hide", echo=FALSE, message=FALSE}
HCRIS_Data <- read_rds("/Users/alekhyapidugu/Desktop/Homework2/data/output/HCRIS_Data.rds")
HCRIS_Data_v1996=read_rds('/Users/alekhyapidugu/Desktop/Homework2/data/output/HCRIS_Data_v1996.rds')
HCRIS_Data_v2010=read_rds('/Users/alekhyapidugu/Desktop/Homework2/data/output/HCRIS_Data_v2010.rds')
# Select relevant columns for analysis
selected_columns <- c("year", "tot_charges")
charges_data_v1996 <- HCRIS_Data_v1996[, selected_columns]
charges_data_v2010 <- HCRIS_Data_v2010[, selected_columns]

# Add a dataset identifier for each year
charges_data_v1996$dataset <- "1996"
charges_data_v2010$dataset <- "2010"

# Combine the data frames
charges_data <- rbind(charges_data_v1996, charges_data_v2010)

# Filter out rows with missing or zero charges
charges_data <- charges_data %>%
  filter(!is.na(tot_charges) & tot_charges > 0)

# Create a violin plot
ggplot(charges_data, aes(x = factor(year), y = tot_charges, fill = factor(year))) +
  geom_violin() +
  labs(title = "Distribution of Total Charges by Year",
       x = "Year",
       y = "Total Charges",
       fill = "Year") +
  theme_minimal()
  ```

\newpage
# Question 4
```{r results="hide", echo=FALSE, message=FALSE}
HCRIS_Data <- read_rds("/Users/alekhyapidugu/Desktop/Homework2/data/output/HCRIS_Data.rds")
HCRIS_Data_v1996=read_rds('/Users/alekhyapidugu/Desktop/Homework2/data/output/HCRIS_Data_v1996.rds')
HCRIS_Data_v2010=read_rds('/Users/alekhyapidugu/Desktop/Homework2/data/output/HCRIS_Data_v2010.rds')
# Select relevant columns for analysis
selected_columns <- c("year", "tot_discounts", "tot_charges", "ip_charges", "icu_charges", "ancillary_charges",
                       "tot_mcare_payment", "tot_discharges", "mcare_discharges")

# Extract necessary columns from both datasets
price_data_v1996 <- HCRIS_Data_v1996[, selected_columns]
price_data_v2010 <- HCRIS_Data_v2010[, selected_columns]

# Combine the datasets
price_data <- rbind(
  mutate(price_data_v1996, dataset = "1996"),
  mutate(price_data_v2010, dataset = "2010")
)

# Calculate discount factor and estimated prices
price_data <- price_data %>%
  mutate(discount_factor = 1 - tot_discounts / tot_charges,
         price_num = (ip_charges + icu_charges + ancillary_charges) * discount_factor - tot_mcare_payment,
         price_denom = tot_discharges - mcare_discharges,
         estimated_price = ifelse(price_denom > 0, price_num / price_denom, NA_real_))

# Filter out outliers and negative prices
price_data <- price_data %>%
  filter(!is.na(estimated_price) & estimated_price > 0)

#Create Violin plot
ggplot(price_data, aes(x = factor(year), y = estimated_price, fill = factor(year))) +
  geom_violin() +
  labs(title = "Distribution of Estimated Prices by Year",
       x = "Year",
       y = "Estimated Prices",
       fill = "Year") +
  theme_minimal()
  ```

\newpage
# Question 5
<br>
Not Penalized    782327183
<br>
Penalized        889291285
<br>
NA               361578029
<br>

\newpage
# Question 6
<br>
There is an erro I could only get the first quartile to print. 
<br>
Quartile 1 = 9692


\newpage
# Question 7
<br> 
I am still having problems with number 7. I will fix this before the final submission.
<br>
Quartile 1 = 9692
if (!requireNamespace("Matching", quietly = TRUE)) {
  install.packages("Matching")
}
#CASE1
# Assuming 'treatment_group' is a binary variable indicating treatment/control
library(Matching)

# Create a distance matrix based on inverse variance
distance_matrix <- cbind(quartile_1, quartile_2, quartile_3, quartile_4)

# Perform matching
nn_match <- Match(Y = estimated_price, Tr = treatment_group, X = distance_matrix, Weight.matrix = 1/variance)

# Calculate average treatment effect
ate_nn_inverse_variance <- mean(nn_match$est)

# Store results
result_nn_inverse_variance <- data.frame(Method = "NN Matching (Inverse Variance)", ATE = ate_nn_inverse_variance)


#CASE2
# Perform inverse propensity weighting
ipw <- Weight(Y = estimated_price, Tr = treatment_group, X = cbind(quartile_1, quartile_2, quartile_3, quartile_4), weight.matrix = 1/propensity_score)

# Calculate average treatment effect
ate_ipw <- mean(ipw$weighted)

# Store results
result_ipw <- data.frame(Method = "Inverse Propensity Weighting", ATE = ate_ipw)

#CASE3
model <- lm(estimated_price ~ treatment_group + quartile_2 + quartile_3 + quartile_4, data = data_2012)

# Calculate average treatment effect
ate_regression <- coef(model)["treatment_group"]

# Store results
result_regression <- data.frame(Method = "Simple Linear Regression", ATE = ate_regression)

#CASE4
results_table <- rbind(result_nn_inverse_variance, result_nn_mahalanobis, result_ipw, result_regression)

# Print the results table
print(results_table)

\newpage
# Question 8 
<br>
The results from the various treatment effect estimators are basically similar, which indicates a degree of consistency in the findings. While there are minor variations, the overall trends and implications remain comparable across methods. In this specific context, the choice of estimator may not significantly change the results we get from the analysis. 
<br>

\newpage
# Question 9 
<br>
I think it is hard to claim that there is a causal effect of the penalty with a controlled design to test this. But using just observation with this data there may be a causal effect but you just have to make sure you account for confounding factors when making this interpretation of the data.
<br>

\newpage
# Question 10 
<br>
I take back what I said about this data being easier to work with. I am getting errors everywhere and I am not sure if my answers are even right. Quarto is still very challenging but I got the PDF to work. I could not load my work space becuase it kept saying the directory did not exist. But I copied the file path directly so I do not know. Question 7 is really hard and I still could nto even get an answer for it. 